# Master Thesis Johanna Schulz

This .Rmd file includes the code for the econometric analysis for the master thesis of Johanna Schulz.
Specifically, this file include the analysis for the main analysis.

## Load libraries
```{r}
install.packages(readxl)
library(readxl)

install.packages(ggplot2)
library(ggplot2)

install.packages(ggcorrplot)
library(ggcorrplot)

install.packages(dplyr)
library(dplyr)

install.packages(lmtest)
library(lmtest)

install.packages(tidyverse)
library(tidyverse)

install.packages(tseries)
library(tseries)

install.packages(car)
library(car)

install.packages("plm")
library(plm)

install.packages("ggtext")
library(ggtext)

install.packages("reshape2")
library(reshape2)
```

## Import and view the data set
```{r}
file_path <- "../outputs/monthly_vwap_brand_CPI_data.xlsx"

data <- read_excel(file_path)
```


```{r}
head(data)
str(data) # I inspect the data types of the variables and will make some changes now#
```
The data types of some of the columns is not perfectly assigned for the following analysis in R. Therefore, some are converted:
```{r}
data$Industry <- as.factor(data$Industry) # because it is a categorical variable
data$Sector <- as.factor(data$Sector) # same applies as above
data$StrongBrand <- as.factor(data$StrongBrand) # dummy variable, therefore factor type is better than numerical
data$MultipleStrongBrands <- as.factor(data$MultipleStrongBrands) # same applies as above
data$Crisis <- as.factor(data$Crisis) # same applies as above
```


## Econometric Analysis

```{r}
# Convert data to a pdata.frame, 'CompanyName' and 'MonthYear' are cross-sectional and time identifiers
pdata <- pdata.frame(data, index = c("CompanyName", "MonthYear"))
```

### Descriptive Analysis

#### Correlation Analysis of the (Continuous) Variables
```{r}
# Correlation between VWAP and other numerical variables
data_corr <- data[, c("VWAP", "BrandRanking", "CPI", "TR", "EBIT", "EBITDA", "VWAP_lagged")]

corr_matrix <- cor(data_corr, use = "complete.obs")

corr_matrix
```


```{r}
library(ggplot2)
library(reshape2)
library(ggcorrplot)
# Compute the correlation matrix
data_corr <- data[, c("VWAP", "BrandRanking", "CPI", "TR", "EBIT", "VWAP_lagged")]
corr_matrix <- cor(data_corr, use = "complete.obs")

# Transform the correlation matrix into a long format
corr_long <- melt(corr_matrix)

# Create a heatmap with correlation values, Times New Roman font, a descriptive title, and no axis titles
ggplot(corr_long, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(value, 3)), color = "black", size = 3, family = "Times New Roman") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  labs(title = "Correlation Heatmap",
       x = "",  
       y = "") +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, family = "Times New Roman"),
        axis.text.y = element_text(family = "Times New Roman"),
        plot.title = element_text(family = "Times New Roman", hjust = 0.5, size = 12),
        plot.subtitle = element_text(family = "Times New Roman"),
        plot.caption = element_text(family = "Times New Roman"),
        legend.title = element_text(family = "Times New Roman"),
        legend.text = element_text(family = "Times New Roman")) + 
  coord_fixed()
```

#### Scatter Plot VWAP - CPI
This graph is only viewed by the author and not included in the thesis.
```{r}
ggplot(data, aes(x = CPI, y = VWAP)) +
  geom_point() +
  labs(title = "Scatter Plot of VWAP vs CPI", x = "CPI", y = "VWAP") +
  theme_minimal()

```
```{r}
# colored by TickerCode
ggplot(data, aes(x = CPI, y = VWAP, color = TickerCode)) +
  geom_point() +
  labs(title = "Scatter Plot of VWAP vs CPI by Company", x = "CPI", y = "VWAP") +
  theme_minimal() +
  theme(legend.position = "bottom") 

```

#### Time Series CPI | Figure 4: Time Series of Consumer Price Index per Month
```{r}
library(ggplot2)

# Convert MonthYear to Date
data$MonthYear <- as.Date(data$MonthYear)

# Create the time series plot with base month lines
cpi_plot <- ggplot(data, aes(x = MonthYear, y = CPI)) +
  geom_line() +
  geom_vline(xintercept = as.Date("2019-12-01"), linetype = "dotted", color = "grey") +
  geom_hline(yintercept = 100, color = "grey") +
  labs(title = "Time Series of CPI per Month", 
       x = "Month-Year", 
       y = "CPI (Base: December 2019 = 100)") +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman"),
        axis.text.x = element_text(angle = 45, hjust = 1, family = "Times New Roman"),
        axis.text.y = element_text(family = "Times New Roman"),
        plot.title = element_text(family = "Times New Roman", hjust = 0.5))

cpi_plot
```
Save the plot in the outputs folder
```{r}
ggsave("CPI_Time_Series.png", plot = cpi_plot, width = 10, height = 6, dpi = 300)
```

#### Grouped Summary Statistics for Categorical Variables
```{r}
# Summary of VWAP by StrongBrand
data %>%
  group_by(StrongBrand) %>%
  summarise(Mean_VWAP = mean(VWAP, na.rm = TRUE),
            Median_VWAP = median(VWAP, na.rm = TRUE),
            SD_VWAP = sd(VWAP, na.rm = TRUE))

```

```{r}
# Summary of VWAP by MutlipleBrands
data %>%
  group_by(MultipleStrongBrands) %>%
  summarise(Mean_VWAP = mean(VWAP, na.rm = TRUE),
            Median_VWAP = median(VWAP, na.rm = TRUE),
            SD_VWAP = sd(VWAP, na.rm = TRUE))

```

```{r}
# Summary of VWAP by Crisis
data %>%
  group_by(Crisis) %>%
  summarise(Mean_VWAP = mean(VWAP, na.rm = TRUE),
            Median_VWAP = median(VWAP, na.rm = TRUE),
            SD_VWAP = sd(VWAP, na.rm = TRUE))
```

```{r}
library(dplyr)

# mean of VWAP for each level of Crisis
mean_vwap_by_crisis <- data %>%
  group_by(Crisis) %>%
  summarise(Mean_VWAP = mean(VWAP, na.rm = TRUE))

print(mean_vwap_by_crisis)
```

##### Table 9: Mean VWAP values by Crisis Status and Brand Strength
```{r}
# mean of VWAP for each combination of Crisis and StrongBrand 
mean_vwap_by_crisis_strongbrand <- data %>%
  group_by(Crisis, StrongBrand) %>%
  summarise(Mean_VWAP = mean(VWAP, na.rm = TRUE))

print(mean_vwap_by_crisis_strongbrand)
```

```{r}
library(dplyr)
# Calculate the count of companies for each StrongBrand category
company_count <- data %>%
  group_by(StrongBrand) %>%
  summarise(Count = n_distinct(CompanyName))

print(company_count)
```

```{r}
# visualisation of the information from above, not used in the thesis
library(dplyr)
library(ggplot2)

mean_vwap_data <- data %>%
  group_by(StrongBrand, Crisis) %>%
  summarise(Mean_VWAP = mean(VWAP, na.rm = TRUE)) %>%
  mutate(Category = factor(paste("StrongBrand =", StrongBrand, "- Crisis =", Crisis),
                           levels = c("StrongBrand = 1 - Crisis = 0", 
                                      "StrongBrand = 1 - Crisis = 1",
                                      "StrongBrand = 0 - Crisis = 0",
                                      "StrongBrand = 0 - Crisis = 1")))

library(ggtext)

custom_labels <- c("Crisis = 0<br><span style='margin-top:10px; display:block; text-align:center;'><strong>Strong Brand = 1</strong></span>", 
                   "", 
                   "Crisis = 0<br><span style='margin-top:10px; display:block; text-align:center;'><strong>Strong Brand = 0</strong></span>", 
                   "")

ggplot(mean_vwap_data, aes(x = Category, y = Mean_VWAP, fill = Category)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.6) +
  scale_fill_grey(start = 0.8, end = 0.5) +
  geom_text(aes(label = round(Mean_VWAP, 2)), vjust = -0.5, position = position_dodge(width = 0.6)) +
  labs(title = "Mean VWAP Values by StrongBrand and Crisis Status",
       x = "",  
       y = "Mean VWAP") +
  scale_x_discrete(labels = custom_labels) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_markdown()) 
```

### Models

All models are first without controls, then stepwise control adding, finally, heteroskedasticity consistent standard errors are used.

#### Hypotheses H1-H2 | Model 1.0 (Crisis as IV) to Model 1.1 (CPI as IV)

#### Model 1.0 | Hypothesis H1, Crisis as IV
```{r}
# Model 1.0 without controls, Crisis as IV
linreg1_0 <- lm(VWAP ~ Crisis, data = data)
summary(linreg1_0)
```
##### Stepwise Controls Adding Model 1.0 | Results are reported in Table 11
```{r}
# Adding 1 control: Sector Information
linreg1_0 <- lm(VWAP ~ Crisis + Sector, data = data)
summary(linreg1_0)
```

```{r}
# Adding 2nd control: TR
linreg1_0 <- lm(VWAP ~ Crisis + Sector + TR, data = data)
summary(linreg1_0)
```

```{r}
# Adding 3rd control: EBIT
linreg1_0_ <- lm(VWAP ~ Crisis + Sector + TR + EBIT, data = data)
summary(linreg1_0)
```

```{r}
# Finally: Adding 4th control: lagged VWAP
# Model 1.0 with controls, Crisis as IV
linreg1_0 <- lm(VWAP ~ Crisis + Sector + TR + EBIT + VWAP_lagged, data = data)
summary(linreg1_0)
```

##### OLS Assumptions | Model 1.0 (with controls) "linreg1_0"
Random sample: observations are iid

1. Nonzero expectation: Zero mean for errors
2. Homoscedasticity: Constant variance for errors
3. Normal distribution of errors
4. No multicollinearity

--> All tests need to be NOT significant to meet assumptions

```{r}
library(lmtest)
# 1. Nonzero expectation: Zero mean for errors -> Resest test
resettest(linreg1_0_) #significant, assumption violated
```

```{r}
# 2. Homoscedasticity: Constant variance for errors -> Goldfeld-Quandt Test
gqtest(linreg1_0_) # not significant, assumption holds
```

```{r}
# 3.1 Normaldistributuon of errors -> Plotting residuals against time
plot(residuals(linreg1_0_) ~ data$MonthYear) # random, assumption holds
```

```{r}
library(tseries)
# 3.2 Errors distributed multivariate normal -> Jarque and Bera (1980)
jarque.bera.test(residuals(linreg1_0_)) # slighlty significant -> VIOLATED
```

```{r}
library(carData)
# 4. No multicollinearity -> VIF values of 5-10 indicate a strong correlation
vif(linreg1_0)
```

```{r}
# Finally, adding Heteroskedasticity Robust Standard Errors to the model to account for the violations
library(lmtest)
library(sandwich)
# Robust t test
coeftest_results <- coeftest(linreg1_0, vcov = vcovHC(linreg1_0, type = "HC0"))

print(coeftest_results)
```

#### Model 1.1 | Hypothesis H2, CPI as IV
```{r}
# Model 1.1 without controls, CPI as IV
linreg1_1 <- lm(VWAP ~ CPI, data = data)
summary(linreg1_1)
```

##### Stepwise Controls Adding Model 1.1 | Results are reported in Table 12
```{r}
# stepwise control adding
linreg1_1 <- lm(VWAP ~ CPI + Sector, data = data)
summary(linreg1_1)
```

```{r}
# stepwise control adding
linreg1_1 <- lm(VWAP ~ CPI + Sector + TR, data = data)
summary(linreg1_1)
```

```{r}
# stepwise control adding
linreg1_1_ <- lm(VWAP ~ CPI + Sector + TR + EBIT, data = data)
summary(linreg1_1_)
```

```{r}
# Model 1.1 with all controls, CPI as IV
linreg1_1 <- lm(VWAP ~ CPI + Sector + TR + EBIT + VWAP_lagged, data = data)
summary(linreg1_1)
```

##### OLS Assumptions | Model 1.1 (with controls) "linreg1_1"
```{r}
# 1. Nonzero expectation: Zero mean for errors -> Resest test
resettest(linreg1_1_) #significant, assumption violated
```

```{r}
# 2. Homoscedasticity: Constant variance for errors -> Goldfeld-Quandt Test
gqtest(linreg1_1_) # not significant, assumption holds
```


```{r}
# 3.1 Normal distribution of errors -> Plotting residuals against time
plot(residuals(linreg1_1_) ~ data$MonthYear) # random, assumption holds
```


```{r}
# 3.2 Errors distributed multivariate normal -> Jarque and Bera (1980)
jarque.bera.test(residuals(linreg1_1_)) # slightly significant -> VIOLATED
```

```{r}
# 4. No (strong) multicollinearity -> 
vif(linreg1_1)
```

```{r}
# Finally, adding Heteroskedasticity Robust Standard Errors to the model to account for the violations

# Robust t test
coeftest_results <- coeftest(linreg1_1, vcov = vcovHC(linreg1_1, type = "HC0"))

print(coeftest_results)
```

#### Hypotheses H3-H4 | Model 2.0 (Crisis as IV and StrongBrand as Moderator) to Model 2.1 (CPI as IV and StrongBrand as Moderator)

#### Model 2.0 - Crisis and StrongBrand
```{r}
# without controls
linreg_2_0 <- lm(VWAP ~ Crisis + StrongBrand + (Crisis*StrongBrand), data = data)
summary(linreg_2_0)
```
##### Stepwise Controls Adding Model 2.0 | Results are reported in Table 13
```{r}
# stepwise control adding
linreg_2_0 <- lm(VWAP ~ Crisis + StrongBrand + (Crisis*StrongBrand) + Sector, data = data)
summary(linreg_2_0)
```

```{r}
# stepwise control adding
linreg_2_0 <- lm(VWAP ~ Crisis + StrongBrand + (Crisis*StrongBrand) + Sector + TR, data = data)
summary(linreg_2_0)
```

```{r}
# stepwise control adding
linreg_2_0_ <- lm(VWAP ~ Crisis + StrongBrand + (Crisis*StrongBrand) + Sector + TR + EBIT, data = data)
summary(linreg_2_0_)
```

```{r}
# with all controls 
linreg_2_0 <- lm(VWAP ~ Crisis + StrongBrand + (Crisis*StrongBrand) + Sector + EBIT + TR + VWAP_lagged, data = data)
summary(linreg_2_0)
```

##### OLS Assumptions | Model 2.0 (with controls) "linreg2_0"

```{r}
# 1. Nonzero expectation: Zero mean for errors -> Resest test
resettest(linreg_2_0_) #significant, assumption violated
```

```{r}
# 2. Homoscedasticity: Constant variance for errors -> Goldfeld-Quandt Test
gqtest(linreg_2_0_) # not significant, assumption holds
```

```{r}
# 3.1 Normal distribtion of errors -> Plotting residuals against time
plot(residuals(linreg_2_0_) ~ data$MonthYear) # random, assumption holds
```

```{r}
# 3.2 Errors distributed multivariate normal -> Jarque and Bera (1980)
jarque.bera.test(residuals(linreg_2_0_)) #significant -> VIOLATED
```

```{r}
# 4. No multicollinearity -> VIF values not be higher than 5
vif(linreg_2_0, type ="predictor")
```

```{r}
# Finally, adding Heteroskedasticity Robust Standard Errors to the model to account for the violations

# Robust t test
coeftest_results <- coeftest(linreg_2_0, vcov = vcovHC(linreg_2_0, type = "HC0"))

print(coeftest_results)
```

#### Model 2.1 - CPI and StrongBrand 
```{r}
# without controls
linreg_2_1 <- lm(VWAP ~ CPI + StrongBrand + (CPI*StrongBrand), data = data)
summary(linreg_2_1)
```
##### Stepwise Controls Adding Model 2.1 | Results are reported in Table 14
```{r}
# stepwise control adding 
linreg_2_1 <- lm(VWAP ~ CPI + StrongBrand + (CPI*StrongBrand) + Sector, data = data)
summary(linreg_2_1)
```

```{r}
# stepwise control adding 
linreg_2_1 <- lm(VWAP ~ CPI + StrongBrand + (CPI*StrongBrand) + Sector + TR , data = data)
summary(linreg_2_1)
```

```{r}
# stepwise control adding 
linreg_2_1_ <- lm(VWAP ~ CPI + StrongBrand + (CPI*StrongBrand) + Sector + TR + EBIT, data = data)
summary(linreg_2_1_)
```

```{r}
# with controls  
linreg_2_1 <- lm(VWAP ~ CPI + StrongBrand + (CPI*StrongBrand) + Sector + TR + EBIT + VWAP_lagged, data = data)
summary(linreg_2_1)
```
##### OLS Assumptions | Model 2.1 (with controls) "linreg2_1"
```{r}
# 1. Nonzero expectation: Zero mean for errors -> Resest test
resettest(linreg_2_1_) #significant, assumption violated
```

```{r}
# 2. Homoscedasticity: Constant variance for errors -> Goldfeld-Quandt Test
gqtest(linreg_2_1_) # not significant, assumption holds
```

```{r}
# 3.1 Normal distribtion of errors -> Plotting residuals against time
plot(residuals(linreg_2_1_) ~ data$MonthYear) # random, assumption holds
```

```{r}
# 3.2 Errors distributed multivariate normal -> Jarque and Bera (1980)
jarque.bera.test(residuals(linreg_2_1_)) #significant -> VIOLATED
```

```{r}
# 4. No multicollinearity 
vif(linreg_2_1, type ="predictor")
```

```{r}
# Finally, adding Heteroskedasticity Robust Standard Errors to the model to account for the violations

# Robust t test
coeftest_results <- coeftest(linreg_2_1, vcov = vcovHC(linreg_2_1, type = "HC0"))

print(coeftest_results)
```

#### Hypotheses H4-H6 | Model 3.0 (Crisis as IV and BrandRanking as Moderator) to Model 3.1 (CPI as IV and BrandRanking as Moderator)

##### Aubset Creation to include only companies with strong brands and test within this subset
```{r}
subset_strongbrand <- subset(data, StrongBrand == 1)

head(subset_strongbrand)
```

#### Model 3.0 - Crisis and BrandRanking
```{r}
# without controls
linreg_3_0 <- lm(VWAP ~ Crisis + BrandRanking + (Crisis*BrandRanking), data = subset_strongbrand)
summary(linreg_3_0)
```
##### Stepwise Controls Adding Model 2.1 | Results are reported in Table 15
```{r}
# stepwise control adding
linreg_3_0 <- lm(VWAP ~ Crisis + BrandRanking + (Crisis*BrandRanking) + Sector, data = subset_strongbrand)
summary(linreg_3_0)
```

```{r}
# stepwise control adding
linreg_3_0 <- lm(VWAP ~ Crisis + BrandRanking + (Crisis*BrandRanking) + Sector + TR, data = subset_strongbrand)
summary(linreg_3_0)
```

```{r}
# stepwise control adding
linreg_3_0_ <- lm(VWAP ~ Crisis + BrandRanking + (Crisis*BrandRanking) + Sector + TR + EBIT, data = subset_strongbrand)
summary(linreg_3_0_)
```

```{r}
# with controls
linreg_3_0 <- lm(VWAP ~ Crisis + BrandRanking + (Crisis*BrandRanking) + Sector + TR + EBIT + VWAP_lagged, data = subset_strongbrand)

summary(linreg_3_0)
```
##### OLS Assumptions | Model 3.0 (with controls) "linreg3_0"
```{r}
# 1. Nonzero expectation: Zero mean for errors -> Resest test
resettest(linreg_3_0_) # significant, violated
```

```{r}
# 2. Homoscedasticity: Constant variance for errors -> Goldfeld-Quandt Test
gqtest(linreg_3_0_) # not significant, assumption holds
```

```{r}
# 3.1 Normal distribtion of errors -> Plotting residuals against time
plot(residuals(linreg_3_0_) ~ subset_strongbrand$MonthYear) # random, assumption holds
```

```{r}
# 3.2 Errors distributed multivariate normal -> Jarque and Bera (1980)
jarque.bera.test(residuals(linreg_3_0_)) #significant -> VIOLATED
```

```{r}
# 4. No multicollinearity 
vif(linreg_3_0, type = "predictor")
```

```{r}
# Finally, adding Heteroskedasticity Robust Standard Errors to the model to account for the violations

# Robust t test
coeftest_results <- coeftest(linreg_3_0, vcov = vcovHC(linreg_3_0, type = "HC0"))

print(coeftest_results)
```

#### Model 3.1 - CPI and BrandRanking
```{r}
# without controls
linreg_3_1 <- lm(VWAP ~ CPI + BrandRanking + (CPI*BrandRanking), data = subset_strongbrand)
summary(linreg_3_1)
```
##### Stepwise Controls Adding Model 2.1 | Results are reported in Table 16
```{r}
# stepwise control adding
linreg_3_1 <- lm(VWAP ~ CPI + BrandRanking + (CPI*BrandRanking) + Sector, data = subset_strongbrand)
summary(linreg_3_1)
```

```{r}
# stepwise control adding
linreg_3_1 <- lm(VWAP ~ CPI + BrandRanking + (CPI*BrandRanking) + Sector + TR, data = subset_strongbrand)
summary(linreg_3_1)
```

```{r}
# stepwise control adding
linreg_3_1_ <- lm(VWAP ~ CPI + BrandRanking + (CPI*BrandRanking) + Sector + TR + EBIT, data = subset_strongbrand)
summary(linreg_3_1_)
```

```{r}
# with controls
linreg_3_1 <- lm(VWAP ~ CPI + BrandRanking + (CPI*BrandRanking) + Sector + TR + EBIT + VWAP_lagged, data = subset_strongbrand)
summary(linreg_3_1)
```
##### OLS Assumptions | Model 3.1 (with controls) "linreg3_1"
```{r}
# 1. Nonzero expectation: Zero mean for errors -> Resest test
resettest(linreg_3_1_) #significant, assumption violated
```

```{r}
# 2. Homoscedasticity: Constant variance for errors -> Goldfeld-Quandt Test
gqtest(linreg_3_1_) # not significant, assumption holds
```

```{r}
# 3.1 Normal distribtion of errors -> Plotting residuals against time
plot(residuals(linreg_3_1_) ~ subset_strongbrand$MonthYear) # random, assumption holds
```

```{r}
# 3.2 Errors distributed multivariate normal -> Jarque and Bera (1980)
jarque.bera.test(residuals(linreg_3_1_)) #significant -> VIOLATED
```

```{r}
# 4. No multicollinearity 
vif(linreg_3_1, type = "predictor")
```

```{r}
# Finally, adding Heteroskedasticity Robust Standard Errors to the model to account for the violations

# Robust t test
coeftest_results <- coeftest(linreg_3_1, vcov = vcovHC(linreg_3_1, type = "HC0"))

print(coeftest_results)
```

### Robustness Tests

```{r}
# with subset for crisis = 0
data_no_crisis <- subset(data, Crisis == 0)

linreg2_no_crisis <- lm(VWAP ~ StrongBrand, data = data_no_crisis)
summary(linreg2_no_crisis)

```



